---
title: "Homework 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, warning = FALSE}
library(gapminder)
library(dplyr)
library(ggplot2)
```


# Task Option 1
_Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year._
```{r}
world_lifeExp <- gapminder %>%
  summarize(mean = mean(lifeExp))
```

```{r, fig.height = 3, fig.width = 15}
gapminder %>%
  select(continent, country, year, lifeExp) %>%
  group_by(lifeExp) %>%
  filter(lifeExp < world_lifeExp) %>%
  ggplot(aes(year, group = country)) +
  geom_bar(width = 0.9) +
  facet_wrap(~continent, nrow = 1) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
gapminder %>%
  select(continent, country, year, lifeExp) %>%
  group_by(continent, year) %>%
  filter(lifeExp < 59.47444) 
  #ggplot(aes(continent, year)) +
  #geom_col(country)

```




# Task Option 2
_Get the maximum and minimum of GDP per capita for all continents._
```{r}

```



# Task Option 5
_How is life expectancy changing over time on different continents?_
```{r}

```

