---
title: "Homework 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, warning = FALSE}
library(gapminder)
library(dplyr)
library(ggplot2)
```


# Task Option 1
__Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.__


_Use summarize() to calculate the grand mean of Life Expectancy for the entire world._
```{r}
world_lifeExp <- gapminder %>%
  summarize(mean = mean(lifeExp))
```

_Then filter out the entries that have a life expectancy that is less than the world mean, and create a tibble to display the data._
```{r, fig.height = 3, fig.width = 15}
lowlifeExp <- gapminder %>%
  select(continent, country, year, lifeExp) %>%
  group_by(lifeExp) %>%
  filter(lifeExp < world_lifeExp)
print(lowlifeExp)
```

_Plot the number of countries in each year for each continent that have a life expectancy that is less than the world mean._
```{r, fig.height = 3, fig.width = 15}
ggplot(lowlifeExp, aes(year, group = country)) +
  geom_bar(width = 0.9) +
  facet_wrap(~continent, nrow = 1) +
  theme(plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = 90)) +
  labs(title = "Count of countries in each year for each continent", x = "Year", y = "Count of countries")
```



# Task Option 2
_Get the maximum and minimum of GDP per capita for all continents, and create a tibble to display the data_
```{r}
minmax_gdp <- gapminder %>%
  group_by(continent) %>%
  summarize(max_gdp = max(gdpPercap),
            min_gdp = min(gdpPercap))
print(minmax_gdp)
```

_Create a boxplot that shows the minimum, maximum, and spread of the data._
```{r}
gapminder %>%
  group_by(continent)%>%
  ggplot(aes(continent, gdpPercap)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "GDP per Capita for each Continent", x = "Continent", y = "GDP per Capita") +
  theme(plot.title = element_text(hjust = 0.5))
```



# Task Option 5
_How is life expectancy changing over time on different continents?_
```{r}

```



















