---
title: "hw 04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(dplyr))
```



# Exercise 1
_Univartiate Option 1_

_1. Make a tibble with one row per year, and columns for life expectancy for two or more countries._
_2. Take advantage of this new data shape to scatterplot life expectancy for one country against that of another._
_3. Re-lengthen the data.__

```{r, widening gapminder data, echo = FALSE}
gap_wide <- gapminder %>%
  select(year, lifeExp, country) %>%
  filter(country == "Canada" | country == "Norway") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              values_from = lifeExp)
gap_wide
```
```{r}
gap_wide %>%
  ggplot(aes(Norway, Canada, fill = year, color = year)) +
  geom_point(size = 3)
```


This plot shows the life expectancy for Norway (x-axis) and Canada (y-axis) between 1952 and 2007. I don't think having the data in this wider format is the best way to format the data in order to plot it efficiently.

```{r, relengthen data}
gap_long <- gap_wide %>%
  pivot_longer(cols = c(Canada, Norway),
               names_to = "country",
               values_to = "lifeExp") %>%
  arrange(country)
gap_long
```

In this format, it will be much easier to display the data efficiently.See below for a plot that is easier to interpret. 

```{r}
gap_long %>%
  ggplot(aes(year, lifeExp, group = country)) +
  geom_line(aes(color = country))
```




# Exercise 2
_Multivariate Option 1_

_1. Make a tibble with one row per year, and columns for life expectancy and GDP per capita (or two other numeric variables) for two of more countries_
_2. Re-lengthen the data._
```{r}
gap_multi_wide <- gapminder %>%
  select(country, year, lifeExp, gdpPercap) %>%
  filter(country == "Canada" | country == "Norway") %>%
  pivot_wider(id_cols = year,
              names_from = country,
              names_sep = "_",
              values_from = c(lifeExp, gdpPercap))
gap_multi_wide
```

```{r}
gap_multi_long <- gap_multi_wide %>%
  pivot_longer(cols = c(lifeExp_Canada, lifeExp_Norway, gdpPercap_Canada, gdpPercap_Norway),
               names_to = c(".value", "country"),
               names_sep = "_",
               values_to = c(lifeExp, gdpPercap)) %>%
  arrange(country)
gap_multi_long
```




# Exercise 3
```{r, echo = FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

## 3.1
_For each guest in the guestlist (guest tibble), add a column for email address, which can be found in the email tibble._

First, I wanted to separate the guests from each family in email - now email, matches the format of the "guest". Then I renamed the "guest" column in email to "name" so that it also matches guest.
Next, I used left_join to add the emails for each person. *note/question: why does it need to be "c("name")" in the left_join function? Why didn't just "name" work (only added email for one person in each party)?
Lastly, I selected for party, name, and email to make the tibble easier to look at. 

```{r, separate guests and rename guest}
email_sep <- 
  email %>% 
  separate_rows(guest, sep = ", ") %>%
  rename(name = guest)

guest2 <- guest %>%
  left_join(email_sep, by = c("name")) %>%
  select(party, name, email)
guest2
```


## 3.2
_Who do we have emails for, yet are not on the guestlist?_
I used anti_join to find the people in the email file who were not included in the guest file. 
```{r}
email_sep %>%
  anti_join(guest, by = c("name"))
```


## 3.3
_Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist)._

```{r}
guest_list <- guest %>%
  full_join
```













